# Explanation Framework Analysis {#explanations}

```{r setupexplanation, echo=FALSE, message=FALSE, warning=FALSE}
source("R/esg_explanations_summary.R")
```

To better understand the nature of data gaps we consulted with data management experts at the World Bank to
develop a list of explanatory factors which are described below. These "explanations" are defined such that they can be used
to consistently "tag" indicators to which they apply. They are particularly relevant to time series statistics.
Using metadata in the World Bank API and interviews with members
of the data management team, we examined each indicator through the lens of these explanations and recorded the results
as dummy variables in a metadata database. It is possible that multiple explanations apply to any given indicator.

The results of this tagging exercise were used to build a database along with other relevant metadata.
The final database assembled for analysis has a total of 48 variables for the 134 indicators.

The discusion below describes each of the explanations in detail. Each description includes a coverage heat
map from the previous section which identifies the indicators "tagged" underneath it, showing how the explanation
may impact gaps in data coverage.

*TGH: Need to replace "Methodology" with "Structural" in all text and charts*

*TGH: need to double check the heatmaps below as the indicator counts don't match the current metadata table. The counts
don't match for "no_gap" and explanations B and E. F & G look right and I can't tell about C. But that's just summary counts
we should really confirm they match at the indicator level. Here are the counts from the database.*

*no_gap    12
expl_a    15
expl_b    15
expl_c    66
expl_d     2
expl_e     8
expl_f    32
expl_g    12*


## Description of Explanations

### Ideal Case {-}

The World Bank API defines 217 countries and economies. Indicators that had observations for 2018 or later (this research
was conducted in 2019) for at least 196 countries (90%) were considered "non gap" indicators and were not tagged with
any further explanations. This category also includes 4 indicators comprised of modeled projections for future time periods.

```{r hm-expl, fig.width = 8, fig.cap = "Ideal Case"}
hm_expl(d1a, ideal, "Ideal case")
```

### Dropped or Deprecated (A) {-}

While most major World Bank time series databases are updated on regular schedules, it is common for individual indicators
and occasionally entire databases to be discontinued. This can occur if the original provider or program discontinues
support for the data, or if policies, programs, or methodologies evolve such that the data are no longer relevant.
In these cases, the time series is no longer updated with recent observations, and the series becomes increasingly obsolete.
Deprecated indicators in the WDI are typically removed and are available only in the [WDI Archives][wdi-archives].

```{r hm-expla, fig.width = 8, fig.cap = "Dropped or Deprecated"}
hm_expl(d1a, expl_a, "Dropped or Deprecated")
```

**Definition:** the indicator only exists in a discontinued or "archived" database, including:

* [WDI Archives][wdi-archives] (API source=2)
* [Africa Development Indicators][adi] (API source=11)

**Notes:** indicators tagged with this explanation are considered inactive and remaining explanations are not considered.


### No Longer Updated (B) {-}

In other cases, an indicator may remain in an active database but fall out of active maintenance. This may be the case as a precursor
to Explanation A above, or it may be attributable to lapsed oversight, i.e., there is no longer a person or team responsible
for the indicator (e.g., program termination, department reorganization, personnel changes). In these cases as well, the time series
becomes increasingly obsolete as it is no longer updated.

```{r hm-explb, fig.width = 8, fig.cap = "No Longer Updated"}
hm_expl(d1a, expl_b, "No Longer Updated")
```

**Definition:** the indicator has no observations for any country for the last five years (2014-present)

**Notes:** indicators tagged with this explanation are considered inactive and remaining explanations are not considered.

### Structural Lags (C) {-}

Some indicators by their nature may be especially time consuming to produce. For instance, if an indicator relies on
administrative data or other underlying raw information, it cannot be calculated until that source is published. These
source dependencies may be subject to their own production schedules and timeliness issues. The effect may be that an indicator
may only be available significantly later than the time period of its observations.
These delays may be compounded if multiple dependencies are involved. 

```{r hm-explc, fig.width = 8, fig.cap = "Methodology Lags"}
hm_expl(d1a, expl_c, "Methodology Lags")
```

**Definition:** classification under this explanation was made through consultations with subject matter experts.

### Curation Lags (D) {-}

In many cases the World Bank is a distributor of indicators it obtains from dozens of external parties.
There is some level of "curation" overhead costs to obtain, collate, validate and finalize data from these independent
sources, and this overhead may itself affect data availability compared to that of the original sources.
This may be especially true where the production cycle of the original indicator varies from that of the Bank's curation team;
for instance, an indicator published monthly by the original source may only be updated quarterly or annually by the World Bank.

```{r hm-expld, fig.width = 8, fig.cap = "Curation Lags"}
hm_expl(d1a, expl_d, "Curation Lags")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.

### Licensing Constraints (E) {-}

In some cases, the World Bank distributes indicators under legal agreements that affect its availability. The most common
case are licenses that stipulate that the Bank cannot distribute the latest version of a dataset in order to protect that
dataset's value in cases where providers sell the data commercially. Thus, while more recent data may exist, the Bank is
legally prevented from distributing it for a period of time, which can constitute a significant availability gap.

```{r hm-exple, fig.width = 8, fig.cap = "Licensing Constraints"}
hm_expl(d1a, expl_e, "Licensing Constraints")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.

### Survey Dependencies (F) {-}

Some indicators are calculated from underlying [microdata](https://stats.oecd.org/glossary/detail.asp?ID=1656) obtained from
household-level or firm-level surveys conducted in country. The survey process constitutes a unique kind of methodological lag.
Surveys themselves are costly, and are frequently implemented irregularly or according to schedules that vary significantly by country.
As a result, data availability may vary not only by time but by country.

```{r hm-explf, fig.width = 8, fig.cap = "Survey Dependencies"}
hm_expl(d1a, expl_f, "Survey Dependencies")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.


### Limited Relevance (G) {-}

Some indicators may not be relevant to or may not be interesting in the context of certain groups of economies. For example,
very small economies may not have significant natural resource endowments or particular kinds of economic activity for particular
indicators to be meaningful. Indicators designed to measure use of, say, forests or mineral deposits, or levels of trade, may
be assumed to essentially be zero or "too small to measure." Similarly, other indicators may not be relevant to industrialized
economies, such as prevalence of certain diseases (thought to be eradicated), literacy rates, or foreign aid.

```{r hm-explg, fig.width = 8, fig.cap = "Limited Relevance"}
hm_expl(d1a, expl_g, "Limited Relevance")
```

**Definition:** this explanation was considered for high-income "rich" countries (using the WBG income classification) and small economies,
defined as those with populations under 120,000. This threshold represents a "natural" breakpoint at which population jumps by nearly
40 percent, clearly differentiating economies above and below it. For both "rich" and "small" economies we identified a set of indicators
for which at least 80% of economies had no observations at all. From these results we removed indicators that we still considered
potentially relevant to rich/small economies despite the lack of available data.


## Findings 

*TGH: I think we should keep the charts as is except possibly 3.11 and write narrative around them, but it would be great if they could
all use the same x axis*


Figure \@ref(fig:smry-exlp) shows the results of the explanation framework 
disaggregated primarily by explanation, and secondarily by primary source 
(WBG and all external sources).  

Overall, the "Methodology Lags" explanation is by far the most dominant cause 
of gaps in ESG data, where it is a contributing factor for 72 of 134 indicators 
overall, and 40 of 67 indicators in the WBG's ESG dataset. This frequency likely
reflects prevalence of traditional statistical indicators used in ESG, many of 
which rely on inter-agency cooperation and other intrinsically time consuming 
modalities to produce comparable data for all countries.

"Survey Dependencies" is the second most significant explanation, contributing 
to gaps for 32 of 134 indicators overall, and 18 of 67 indicators in the WBG's 
dataset. Most of these indicators (75% and 83% respectively) fall under the 
"Social" pillar (See Figure \@ref(fig:smry-exlp-sctr)).

```{r smry-exlp, fig.width = 12, fig.height = 5.5, fig.cap = "Explanations for data gaps"}
smry_expl <- smry_expl +
  labs(
    title = "Explanations for gaps: All indicators vs WBG dataset",
    subtitle = "Most data gaps are due to methodological lag and lack of survey data. 
    This is true both when considering all indicators, or the subset of indicators
    contained in the WBG dataset.",
    y = "Frequency",
    x = "",
    fill = "Data producer"
  )

plotly::ggplotly(smry_expl, tooltip = "text") %>% plotly::style(hoveron = "fill")

```


```{r smry-exlp-sctr, fig.width = 12, fig.height = 5.5, fig.cap = "Explanations for data gaps, by sector"}
smry_expl_sctr <- smry_expl_sctr +
  labs(
    title = "Explanations for gaps by sector: All indicators vs WBG dataset",
    subtitle = "Not only most data gaps are due to methodological lag, it is one of
    the most frequent explanation for all three ESG sectors.",
    y = "Frequency",
    x = "",
    fill = "Sector"
  )

plotly::ggplotly(smry_expl_sctr, tooltip = "text") %>% plotly::style(hoveron = "fill")
```

**ANOTHER WAY TO LOOK AT THE SAME INFORMATION AS CHART ABOVE**
**I like this one because it is easy to focus what is going on sector by sector**
**I thought there would be double counting, but there is not. Your logic probably took care of this**
```{r smry-exlp-sctr2, fig.width = 12, fig.height = 5.5, fig.cap = "Explanations for data gaps, by sector"}
smry_expl_sctr2 <- smry_expl_sctr2 +
  labs(
    title = "Explanations for gaps by sector: All indicators vs WBG dataset",
    subtitle = "Not only most data gaps are due to methodological lag, it is one of
    the most frequent explanation for all three ESG sectors.",
    y = "Frequency",
    x = "",
    fill = "Explanation"
  )

plotly::ggplotly(smry_expl_sctr2, tooltip = "text") %>% plotly::style(hoveron = "fill")
```


```{r shr-ext, fig.width = 10, fig.height = 5, fig.cap = "Share of externally-sourced indicators"}
shr_ext <- shr_ext +
 labs(
    title = "Share of externally sourced indicators by explanation",
    x = "",
    y = ""
  )

plotly::ggplotly(shr_ext, tooltip = "text")
```


**ORIGINAL PARAGRAPH STARTS HERE**
*Figure 3.1* shows the results of the explanation framework disaggregated
primarily by explanation, and secondarily by cluster (E, S and G) and
primary source (WBG and all external sources).  Table X reflects the
entire universe of indicators studied (134 indicators total) while Table Y includes only those
in the current version of the WBG ESG-specific dataset (67 indicators total).

Overall, the "Methodology Lags" explanation is by far the dominant cause of gaps in ESG data,
where it is a contributing factor for 72 of 134 indicators overall, and 40 of 67 indicators in the
WBG's ESG dataset. This frequency likely reflects prevalence of traditional statistical indicators
used in ESG, many of which rely on inter-agency cooperation and other intrinsically time consuming
modalities to produce comparable data for all countries.

"Survey Dependencies" is the second most significant explanation, contributing to gaps for 32 of
134 indicators overall, and 18 of 67 indicators in the WBG's dataset. Most of these indicators
(75% and 83% respectively) fall under the "Social" pillar.

15 of 134 indicators overall have been dropped or deprecated by the World Bank and are no longer
maintained. Fortunately, the WBG's dataset does not contain any deprecated indicators. An
additional 11 indicators overall of which 4 are the WBG dataset have not been updated in
at least 5 years, suggesting that the underlying program may have been discontinued in some
fashion. Alternatively, some of these indicators may actually reflect an especially long "methodology
lags"; for instance, indicators produced on 5- or even 10-year intervals.

8 of 134 indicators overall and 6 of 67 indicators in the WBG dataset are subject to licensing
restrictions which may adversely impact the availability of recent data. Only 2 indicators are
affected by differences in the curation cycle, in which the WBG updates its data less frequently
than does the originating source.

Finally, 12 of 134 indicators overall and 6 of 67 indicators in the WBG dataset may by
definition not be relevant to high-income or very small economies. This explanation is a
somewhat exceptional case, as it may suggest special treatment at the analytical level
as opposed to strategies to improve country coverage in the data.

### Results by Source and Cluster {-}

*Figures 3.2 and 3.2* show results of the explanation framework in percentage terms, disaggregated
by primary source and cluster respectively.
Overall, just over 73% of ESG indicators studied originate from sources other than the World Bank.
Figure X shows how this factor varies by explanation. Not surprisingly, all indicators restricted
by licenses come from external sources. On the other hand, the two explanations with the lowest
share of externally sourced indicators (and thus the highest share of World Bank sourced indicators)
are 'Curation Lags' (which only includes 2 indicators) and 'Survey Dependencies.' This is significant
because it may be easier for the World Bank to improve data coverage in areas where it is the primary
source for a significant portion of data.

Results vary significantly by cluster and explanation. All indicators classified as 'Deprecated' along
with all indicators subject to Licensing restraints are in the Environmental cluster, whereas both
indicators in the 'Curation Lags' explanation are in the Social cluster. 'Survey Dependencies' is also heavily
weighted with indicators in the Social cluster. The remaining explanations are distributed 
among clusters in roughly line with their respective portions of the overall sample.



[wdi-archives]: https://datacatalog.worldbank.org/dataset/wdi-database-archives
[adi]: https://datacatalog.worldbank.org/dataset/africa-development-indicators
