# Coverage Analysis {#coverage}

```{r setupcoverage, echo=FALSE, message=FALSE, warning=FALSE}
#--------- execute script of analysis
source("R/load_data.R")
source("R/esg_availability.R")
```


The availability of ESG indicators by country is dynamic over time. As shown in Figure \@ref(fig:ciavailability), not all indicators are available in all countries neither they are available in all years. The availability of indicators goes from darker to lighter. The lighter the color the more the greater the number of countries per indicator in each year. Several insights are drawn from this chart. 

```{r ciavailability, fig.width = 12, fig.height = 10, fig.cap = "Number of countries per indicator over time"}
pg1
```

The characteristics of ESG indicators coverage vary significantly across the study area, 
but can be classified under one of five different patterns:

*TGH: a couple of overall observations on 2.1 - 2.5. 1) the study period has changed (now starts at 2000) and the text hasn't really been updated to reflect
that. Need to accordingly review and revise summary stats as well as description in 2.1; 2) I'm not sure 2.2 is relevant any more; 3) 2.4
is the most important category since this is how we define "data gap" in the introduction, so let's underscore that point; 4) can we
add a complete list of indicators for each section to help quantify the analysis? 5) 2.5 (high coverage) corresponds to "no gaps" classification
in the next section so let's make that point and make sure the indicators align.*


## Low coverage

```{r echo = FALSE}
tr_low <- 100  # number of countries. 
d3 <- d2  %>%
  mutate(max = max(mean),
         less = (mean < tr_low)) %>%
  summarise(sless = sum(less),
            max = mean(max),
            s   = sless/max)
d4 <- d2 %>% 
  filter(mean < tr_low)


```

Some indicators have very low availability over time. In general, `r d3[["sless"]]` (`r round(100*d3[["s"]],2)`%) indicators are available for less than `r tr_low` countries in average over time. Yet, not all indicators with low coverage have the same characteristics. For instance, 

```{r dtlc, options}
d4 %>% 
  arrange(mean) %>% 
  DT::datatable(
    colnames = c("Code", "Indicator", "Average of countries per year"),
    rownames = FALSE,
    filter = 'top', 
    width = '55%', 
    extensions = 'FixedColumns',
    options = list(pageLength = 5)
  ) %>% 
  DT::formatRound("mean" , 1)

```


## Remarkable improvement over time

There are a few indicators that show a remarkable improvement in country coverage over time. Despite their early appearance in the beginning of the 1980's, indicators that come from household surveys such as the _Poverty headcounts_ or _Gini index_ have shown a dramatic improvement in country coverage in the beginning of the 2000's.  The coverage improved from about five countries in the 80's to more than 70 countries per year in the decade of the 2000's.  However, these indicators suffer from a compilation lag that requires more than two years from time or recollection to the time of publication. 
*TGH: again, the comparison to the 1980's isn't really relevant anymore*

Other indicators, especially the ones related to the governance component.  have improved the frequency of publication. For instance, _Governance effectiveness_, _Political stability and absence of violence_, and _Voice and accountability_ were collected every three years at the decade of 1990, but increased their frequency to every year since the 2000. 

## Remain stable over time but with middle coverage

All the indicators with a stable yellow  vertical line are characterized by being stable over time but with a country coverage of about 120 countries per year. Among those, we could find _Prevalence of HIV in adult population_, _non-residents patents application_, or _Physicians per 1,000 people_. 
*TGH: the color map changed, so these are no longer yellow. I'm not sure what it is: burgundy perhaps?*

## High coverage and sudden decline

With the exception of a few indicators like _Ease of doing business index_ or _Total tax and contribution rate_, most recent value (MRV) of most indicators is two or more years old. However, some indicators with remarkable good coverage in the previous decade suddenly declined in the last five years. For instance, _CO2 emissions_ or _Renewable electricity output_ had a yearly coverage of more than 180 countries until 2014/15 and thereafter it suddenly declined to one country, Sudan, for _CO2 emissions_ or zero countries for _Renewable electricity output_. A similar story happened to _Net energy imports_, _Electricity production from nuclear sources_, or _Hospital beds_. However, this set of indicators did not have such a good coverage as _CO2 emissions_ before the decline. 

## high coverage 

Finally, the indicators at the right side of the chart present the best country coverage over time. They are characterized by their intense blue color. For instance, _Population density_, _Adjusted savings: mineral depletion_, or _Land area_ have a remarkable and stable high country coverage. 


