# Coverage Analysis {#coverage}

```{r setupcoverage, cache=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
#--------- execute script of analysis
source("R/load_data.R")
source("R/esg_availability.R")
```

While the 2018 paper looked at coverage gaps primarily in terms of most
recent available values (MRVs), in this analysis we wanted to develop a
more detailed approach to identify different types of coverage gaps over
a broader time span. Accordingly, we developed the heat map shown in
Figure \@ref(fig:ciavailability). In this chart, discrete indicators are
arranged along the Y axis while time is plotted on the X axis for the
2000-2018 period. Colors indicate the number of observations (i.e.,
countries) for the corresponding indicator and year. Darker colors in
the purple part of the spectrum indicate relatively low-density
coverage, while lighter colors in the yellow part of the spectrum
represent high-density coverage, up to the maximum of 217 countries.
Blank areas indicate no data for that particular indicator and year.

```{r ciavailability, fig.width = 12, fig.height = 10, fig.cap = "Number of countries per indicator over time"}

report.graphic(g1, 'ciavailability',
    pdf=list(
      layers=list(
        labs(y="Indicators"),
        theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          axis.text.x = element_text(size=rel(0.5)),
          axis.title.y=element_text(size=rel(0.5)),
          plot.margin=margin(t=20))
      ),
      args=list(width=4, height=5)
    ))

```

Several patterns emerge from a visual assessment of Figure
\@ref(fig:ciavailability), which are not necessarily mutually exclusive,
nor is a visual assessment the only approach to identifying indicator
clusters. [Appendix 2](appendix-coverage-analysis.html) includes several
alternatives to and in-depth definitions of the patterns discussed in
this section.

## Consistently low coverage

```{r echo = FALSE}
tr_low <- 100  # number of countries.

d0 <- x %>%
  group_by(indicatorID,indicator,date) %>%
  summarise(series_date_count=n()) %>%
  summarise(max_count=max(series_date_count))

d3 <- d0 %>% 
  filter(max_count<tr_low)

```

One group of indicators is characterized by consistently low country
coverage over the 2000-2018 time period. In this case we've defined "low
coverage" has having values for no more than `r tr_low` countries in any
given year. These indicators generally appear as steady and consistently
dark horizontal lines towards the bottom of Figure
\@ref(fig:ciavailability).


```{r lowcov}

dd3 <- d3 %>% 
  select(indicator,indicatorID, max_count) %>% 
  arrange(max_count)

report.table(dd3,
             c('Indicator', 'Code', 'Max. Countries'),
             "Indicators with consistently low coverage",
             html=list(widths=list(20)),
             pdf=list(widths=list(15, 9, 2)))

```


It's important to note, however, that while total country coverage may
be low for these indicators in any given year, the country composition
often varies from year to year for reasons discussed in the next
section. For example, "Poverty Headcount Ratio" is available for no more
than `r d3[d3$indicatorID=="SI.POV.NAHC", 'max_count']` countries in any
given year, but includes values for
`r nrow(x %>% filter(indicatorID=="SI.POV.NAHC") %>% group_by(iso3c) %>% summarise(n()))`
countries across all years. By comparison, "Incidence of Malaria" is
available for `r d3[d3$indicatorID=="SH.MLR.INCD.P3", 'max_count']`
countries in nearly all years with very little variation in any given
year. These patterns may be important for ESG analysis if it is possible
to extrapolate or impute missing values from prior years; indicators
whose countries vary from year-to-year (and thus have larger coverage in
the aggregate) may benefit to a greater degree.

## Moderate to high coverage

By contrast, most indicators include values for at least 100 countries
in at least one year. `r nrow(d0 %>% filter(max_count>=100))` indicators
have single-year coverage of at least 100 countries,
`r nrow(d0 %>% filter(max_count>=150))` indicators cover at least 150
countries, and `r nrow(d0 %>% filter(max_count>=200))` indicators cover
at least 200 countries. In Figure \@ref(fig:ciavailability) these
indicators range from magenta to light yellow in the middle to upper
sections of the heat map.

`r nrow(filter(mtd, no_gap==TRUE))` indicators in this cluster include
values for 2018 or later for at least 90% of countries. These indicators
appear at the top-most section of Figure \@ref(fig:ciavailability) and
correspond to the "perfect case" classification in the next chapter.

Among the remaining indicators, year-to-year composition of coverage can
vary in a manner similar to those in the "consistently low coverage"
group for methodological reasons, as discussed in the next section.

## Measurable improvement

A handful of indicators demonstrate significant, measureable improvement
in country coverage over time. We define "measurable improvement" by
regressing country coverage over time for each indicator (gap-filling
for years where coverage is missing entirely). Indicators with a
coefficient greater than 1 are shown in Table \@ref(tab:tabimproved). In
Figure \@ref(fig:ciavailability) these appear as indicators that are
colored dark purple or magenta on the left side of their coverage with
increasingly light colors on the right side.

```{r tabimproved}
ex <- c('IT.NET.BBND.P2',
'ENF.CONT.COEN.COST.ZS',
'IC.REG.DURS',
'IC.TAX.TOTL.CP.ZS',
'SE.ADT.LITR.ZS',
'IP.PAT.NRES',
'EG.ELC.ACCS.ZS',
'IP.PAT.RESD',
'GFDD.DM.06',
'SG.GEN.PARL.ZS',
'SI.SPR.PCAP.ZG',
'BN.CAB.XOKA.GD.ZS')


timpx <- x %>% 
  select(indicator,indicatorID) %>% 
  distinct() %>% 
  filter(indicatorID %in% ex)

report.table(timpx, c('Indicator', 'Code'),
             "Select indicators that improve over time",
             pdf=list(widths=list(15, 12)))

```

These and similar indicators may warrant further study to better
understand the factors behind the increases in country coverage. For
instance, if country coverage improved as a result of better
methodologies, increased production capacity, or broader demand, they
may provide a model for improving country coverage for indicators that
need it.

## High coverage and sudden decline

```{r echo=FALSE}
d1 <- x %>% 
  group_by(indicatorID) %>% 
  summarise(max_year=max(date)) %>% 
  arrange(max_year) 

```

A significant group of indicators has consistent coverage through most
of the time period, but with declining coverage or no coverage in recent
years. In Figure \@ref(fig:ciavailability) these tend to appear as
"truncated" series with no coloring for large portions of the right side
of the chart. Table \@ref(tab:mrv) summarizes indicators by the the year
of their most recent available value (MRV). As shown, over 50% of ESG
indicators in the study period have no values for the most recent study
year, and 13% of indicators have no values for the most recent four
years or more.

```{r mrv}
columns <- c('2018+', '2017', '2016', '2015', '<2015')
values <- c(nrow(d1 %>% filter(max_year>=2018)),
            nrow(d1 %>% filter(max_year==2017)),
            nrow(d1 %>% filter(max_year==2016)),
            nrow(d1 %>% filter(max_year==2015)),
            nrow(d1 %>% filter(max_year<2015)))

report.table(data.frame(columns, values), c('Year of MRV', '# Indicators'),
             "Indicators by Most Recent Available Year")

```

As noted previously, MRV years are a significant factor in ESG data use,
as older data is less relevant to investment decisions being made today
and in the near future. Many important factors could explain the wide
variance in MRV years, and this is the focus of the next chapter.

## Intermittent coverage

A handful of indicators are only available for periodic years with no
values available for intermediate years. These appear in Figure
\@ref(fig:ciavailability) as as intermittent series resembling "dashed"
lines, the majority (but not all) of which are environmental indicators.
Appendix 2 provides a technical description of indicators in this
category.

```{r tabintermittent}
ex <-
  c(
    'SM.POP.NETM',
    'SH.DTH.COMM.ZS',
    'ER.H2O.INTR.K3',
    'ER.H2O.INTR.PC',
    'NW.NCA.SACO.TO',
    'NW.NCA.SAOI.TO',
    'NW.NCA.SAGA.TO',
    'SH.STA.DIAB.ZS',
    'ER.H2O.FWTL.ZS',
    'EN.POP.EL5M.ZS',
    'EN.MAM.THRD.NO',
    'EN.CLC.MDAT.ZS',
    'WBL'
  )

tix <- 
    x %>% 
    select(indicator,indicatorID) %>% 
    distinct() %>% 
    filter(indicatorID %in% ex) 

report.table(tix,
             c('Indicator', 'Code'),
             "Indicators with intermittent coverage",
             pdf=list(widths=list(20)))
 
```

While not the primary focus of this analysis, there are a handful of
factors that could explain the coverage characteristics of this group.
The most obvious explanation is that indicators may simply not be
designed as time-series data. This is the most likely explanation for
Retirement Age and Threatened Mammal Species, which are available for
only a single year. In other cases, there may not be resources to
collect data on an annual basis, even if doing so would be useful. Other
indicators may measure environmental or social phenomena that change
gradually so that annual data collection would not be efficient. This
last possibility is material to ESG data use because it implies that
older data may still be relevant if properly understood.
