---
title: "ESG indicators"
subtitle: "An Analysis of availability"
author: ""
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
    number_sections: yes
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: 
link-citations: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
pkg <-
  c(
    "tidyverse",
    "data.table", 
    "ggplot2",
    "treemapify", 
    "magrittr", 
    "tufte"
  )
new.pkg <-
  pkg[!(pkg %in% installed.packages()[, "Package"])] # check installed packages
load.pkg <-
  pkg[!(pkg %in% loadedNamespaces())]              # check loaded packages

if (length(new.pkg)) {
  install.packages(new.pkg)     # Install missing packages
}

if (length(load.pkg)) {
  lapp <- lapply(load.pkg, library, character.only = TRUE) # load all packages
}

knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

#knitr::opts_knit$set(root.dir = "D:/R working directory/")
```


```{r loadData, echo=FALSE, message=FALSE, warning=FALSE}
esg2 <- read.csv("data/ESG.csv")
esg <- read_csv("data/ESG.csv")
esg <- esg %>% rename_at(vars(contains(" ")),
                  ~str_replace(., " ", "_")) %>% 
  rename_all(tolower) 

#names(esg) %<>% tolower

```


## Executive Summary {-}

TBD


# Introduction

The current paper is the continuation of the work conducted and resumed in a discussion paper titled “Options for Improving Use of ESG data for Sovereign Bond Analysis”. Following interaction with investors in sovereign bonds that used Environmental, Social and Governance indicators in their country analyses and risk/return profiles of sovereign securities, the World Bank set out to present a set of options for improving accessibility, quality (e.g. timeliness and regularity of publication, geographic coverage) and transparency of World Bank Group (WBG) Emerging Markets data. More specifically of interest were Environmental, Social and Governance data. 
The 2018 paper assessed a series of three options for improving the accessibility and quality of ESG data and selected a total number of 138 indicators for subsequent analysis in this regard. The previous paper found that financial institutions use ESG data either coming from data providers which gather information and sell it to interested parties (e.g. Sustainalytics, Bloomberg, Versik Maplecroft, MSCI) and rating agencies, index providers and asset management firms that include ESG data into their proprietary analytical frameworks or offer ESG investment products (e.g. RobecoSAM, JPMorgan, Moody’s, Thomson Reuters, FTSE Russell, Arabesque and Neuberger Berman). A quick analysis of ESG tools provided by  Bloomberg, MSCI, Robescam and Maplecroft revealed that at least 137 indicators used in these tools are sourced from various WBG databases, such as World Development Indicators (WDI), Doing Business, World Governance Indicators (WGI), Women Business & Law and Global Financial Development among others. Bloomberg used 46 WBG indicators out of a total of 54 ESG indicators. MSCI used 76 WBG indicators out of the total of 131 ESG indicators while Robescam used a total of 8 WBG indicators of a total of 24 ESG indicators. The 2018 paper revealed that a lack of dedicated framework on measuring and understanding ESG approaches and how indicators are used for ESG purposes is a significant barrier to the development of their use by the broader finance community. Additionally, it was revealed that data coverage represented a significant issue in WBG data used for ESG because there are a significant number of cases where data was either out of date by several years or where the most recent value (MRV) varied substantially across countries. The previous analysis showed that the number of ESG indicators for which  recent values are available (2015, 2016, 2017 or later) decreases. 45% of indicators have recent values (2016 or later) for at least 75% of countries in the WBG system. One extreme example is India for where the most recent available poverty and inequality estimate is more than six years old, based on a survey from 2011-2012. At times, older values are carried forward to more recent years by the ESG community in their efforts to bridge the gaps, but problems occur when new data comes in and differs from carried forward data, an occurrence that is hard to explain, more so in dynamic economies where this is frequently the case. 
Having identified these issues in the previous paper, a series of three options have been proposed to improve the use of WBG indicators for ESG activities. The first option was to improve the availability and organization of ESG data currently in use mainly by compiling a dataset of indicators used for ESG dataset, ideally in an online ESG portal following the similar examples of for health, gender, jobs, SDG and other. Option two proposed the exploration of means to fill gaps in existing ESG and option three explored leveraging WBG expertise to provide direction for ESG data and standards. 
The current paper explores the first two options proposed in the 2018 paper. It aims at providing an overview of the indicators curated and disseminated by World Bank and relevant for ESG (Environmental, Social, Governance) investment. The work has been conducted assembling a metadata database on indicators, doing a series of interviews with data managers and analyzing and researching the set of indicators and their metadata. We selected138 indicators of relevance to ESG investments, selection informed by the research and consultations conducted for the 2018 paper and we provide in the following parts of the paper an analytical overview of these indicators. We start first with a detailed overview of the methodology used to analyze the dataset and we continue with providing a set of insights on data availability, sources, processing entities and dissemination of these indicators and the frequency of their updates and release. The first part of the paper aims at providing an overview of the methodology used for the analysis and a description of the database assembled for the analysis. The second part of the paper aims at providing analytical insights on the ESG indicators with a focus on measuring data quality and availability, to improve availability, fill in gaps and organize ESG data under a common framework. 

# Methodology

We assembled a database for analysis consisting of metadata with a total of 34 variables These variables were meant to provide the most comprehensive overview of each indicator separately. In addition to an individual tracking code assigned to each indicator, we categorized them by characteristics such as the database of which the indicators is part of, its topic, its statistical methodology used for compilation, its aggregation methods, its relevance to development, its periodicity, among others. Out of these 34 variables, we identified a set of 17 variables that contain most of the information that is relevant for the analysis of these indicators in the context of their use in ESG investments. This subset of variables has been set out to capture the collecting, processing and dissemination entity, the granularity of data (monthly, yearly etc), and the shortest gap between release date and most recent data point available. Finally, we tried to capture the statistical methods through which the indicators are created and to capture imputations performed on the dataset. The annex of this paper provides an overview of the initial database of the indicators metadata and the short version with the variables of most interest to this paper.  In order to fill in knowledge gaps in metadata, we conducted eight interviews with internal Point of Consultation (PoC), which were selected due to their extensive knowledge of each indicator of interest. The interviews took place over a three week period, in person and over phone and focused mostly on timeliness of updates. 
Environmental indicators are related to climate change, carbon emissions, pollution, resource efficiency and biodiversity. Social indicators pertain to human rights, labor standards, health and safety, diversity policies, development of human capital (health & education). Finally, governance indicators refer to corporate governance issues, corruption, rule of law, institutional strength and transparency. 
Finally, using the database assembled on the abovementioned guidelines we were able to analyze the metadata of indicators and provide insights on key issues such as update frequency and dissemination, processing methods, data handling techniques, and quality of indicators.  Additionally, an analysis of data variability and of data points availability was conducted in order to assess first the impact of changes in indicator values and how old data was. 

# Database description

The final database assembled for analysis had a total of 35 variables for a total of 138 indicators. 47 indicators are environmental indicators, 25 governance indicators and 66 social indicators. The World Bank was is a source for a total of 26 indicators, of which 10 each are environmental and social indicators respectively and 6 are governance indicators. Organizations pertaining to the UN system are a source for 53 indicators of which 41 social indicators, 8 environmental indicators and 4 governance indicators. 57 indicators are sourced from different organization grouped under the term of other organization. 29 of these indicators are environmental, 15 are social  and 13 are governance indicators. 

Source	Env 	Soc	Gov
WBG	10	10	6
UN System	8	41	4
Other org	29	15	13


# Discontinued Indicators

Following the interviews with the data managers, we identified a total of 41 out of 138 indicators that are discontinued. 27 of these indicators are environmental indicators, 9 are governance indicators and 5 are social indicators. Given the number of discontinued indicators, there are a total of 97 ESG indicators that are continuously updated. A total of 20 environmental indicators are left from an initial batch of 47 indicators. This is by far the highest rate of attrition observed in the indicators as 57% of the initial set of indicators appears to be no longer updated. 16 governance indicators are left of a total of 25 original ones representing a 36% attrition rate. Finally, social indicators appear to be the most robust class of indicators with 61 out of 66 original ones remaining equivalent of a 8% attrition rate. 

# Data Collection, Processing and Dissemination

We designed three separate source types to capture variation between entities that collect the data, entities that process and entities that disseminate the datasets. We defined a collecting entity an entity that gathers the raw data through different statistical instruments and methods. A processing entity is defined as an entity that takes raw data and transforms it into indicators and numbers that are meaningful and can be released to a wider audience. Finally, a dissemination entity is defined as an entity that takes data ready for dissemination with wider audiences and disseminates it through their own channels such as online databases, brochures, booklets and other means. 
	5 out of 97 ESG indicators have a different processing entity as compared to the collecting entity. 3 of these indicators are environmental indicators and 2 are social indicators. For example, energy intensity level is calculated per unit of GDP at Power Purchase Parity (PPP) by taking data from IEA on energy indicators and data on GDP from World Bank. There are three indicators only that have a different dissemination entity as compared with the processing entity. Two of these indicators are social indicators and one is a governance indicator. All these indicators are collected and processed by OECD, but disseminated by UNESCO. Finally, a total of 8 indicators have a different dissemination entity as compared to the collection entity. 4 of these are social indicators, 3 are environmental indicators and one is a governance indicators. 5 of these indicators are disseminated by World Bank and 3 by UNESCO, while three of them are collected by IEA, another 3 by OECD and one each by FAO and UN. 
In conclusion, there is little variation between collecting, processing and disseminating entities as organizations usually disseminate through their own channels the datasets that they collect and process. 

# Highest Frequency Production and Updates

There is insignificant variation between the production update by the organizations producing the indicators and the timeline of update at the World Bank. Two indicators only, both social, come with different update timeline at the World Bank as compared to the production update. One of the indicators, Agriculture, forestry, and fishing, value added (% of GDP), is updated quarterly by the collecting entity while the World Bank updates it every six months and another one, Proportion of seats held by women in national parliaments (%), is updated monthly by the collecting entity while the World Bank updates it annually. 
	12 out of 20 environmental indicators are produced yearly by the originator and updated yearly by the World Bank. Four environmental indicators are produced and updated every other year while three are produced and updated every three years. 50 out of 61 social indicators are produced and updated on a yearly basis while 13 out of 16 governance indicators are produced yearly. In total, 75 out of 97 indicators are produced and updated, by the World Bank, on a yearly basis making this both the highest production and update frequency.

# Country Coverage and MRV

ENV
Country coverage of indicators point to the comprehensiveness and availability of data of the analyzed indicators. There are only two environmental indicators that do not have a single country data point in 2010, while there are 6 indicators that have no data point in 2014. As we come to more recent years, lack of data at country level, is more prevalent. There are 18 indicators with not a single data point at country level for 2016 while 3 environmental indicators have data for a single country in 2016. Another 10 indicators have data points for around 196 countries each, on average. Country coverage decreases dramatically in 2018. 28 indicators have zero country coverage for 2018 while three only have one country data point. A number of four indicators only have have on average 211 country datapoints for 2018. 
MRV ? (MIN? MAX ?)

# Variation Coefficient
(Gloss on work done by Andres)

# Other aspects 

67 indicators selected for the ESG portal vs 138 for this analysis ? 


