# Volatility Analysis {#volatility}

```{r setupvariability, echo=FALSE, message=FALSE, warning=FALSE}
#--------- execute script of analysis
source("R/load_data.R")
source("R/esg_variability.R")
```

The previous two chapters sought to provide insights into the patterns of coverage gaps found in
ESG data, and the reasons why those gaps occur from a data production standpoint. This chapter
starts to explore the potential for correcting certain gaps using statistical methods. For instance,
missing values might (within limits) be interpolated or extrapolated from existing observations,
or estimated using linear regression or other modelling techniques. While these techniques are
an imperfect substitute for observed data, they may be able to play a limited role within a broader strategy for improving ESG data coverage.

Note that the scope of this paper is limited to analysis of the *potential* impact of gap filling
and does not look in detail at specific techniques or tools for doing so. In all likelihood, the most appropriate approach will vary by indicator, and will depend on the definition fo the indicator,
the availability of proxies and independent variables (for modelling) and other indicator-specific factors. 

In general though, statistical techniques work best when indicator values move according to observable
patterns, and in particular, when temporal variance is relatively low. That is, the less that an indicator
varies from period to period for a particular country, the easier it is to reasonably estimate missing values.
Consequently, indicator variance can be used as a proxy of the general potential to fill gaps using
statistical methods, with low variance suggesting greater potential. In this analysis, we used a
standard *cofficient of variation* (standard deviation divided by mean) to estimate variance for each
indicator and country, and scaled all values to a 0-1 range to make values comparable across indicators.

Figure \@ref(fig:civar) shows the coefficients of variation (CV) for each indicator and country displayed
as an interactive heat map with indicators on the vertical axis and countries on the horizontal axis.
Relatively low CV values appear in light yellow, while high values appear in dark purple.

Note that indicator/country series consisting of a single year are not included in Figure \@ref(fig:civar)
since there is no "variation" in a domain of just one observation.

```{r civar, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Coefficient of variation by country and indicator", fig.fullwidth = TRUE, fig.width = 14, fig.height = 8}

report.graphic(g2, "civariability",
  pdf=list(
      layers=list(
        labs(x="Countries", y="Indicators"),
        theme(axis.text.y=element_blank(),
            axis.ticks.y=element_blank(),
            axis.title.x=element_text(size=rel(0.5)),
            axis.title.y=element_text(size=rel(0.5)),
            plot.margin=margin(t=20))
      ),
      args=list(width=4, height=5))
)
```



Figure \@ref(fig:civar) is a very dense display of data, but its advantage is that it makes it easy
to spot outlying instances of country/indicator combinations that are relatively volatile over time,
i.e., the dark points with high CV values. These are cases where statistical techniques would be likely be
especially unreliable and should be avoided. In the majority of cases, however, ESG indicators appear consistently stable over time.

Table \@ref(tab:esg-var-percs) shows the percentile thresholds for the CV values in Figure \@ref(fig:esg-var). Figure \@ref(fig:esg-var-percs2) shows the same data in a histogram; the red
line marks the median, while the blue lines mark the 5th and 95th percentiles.

```{r esg-var-percs}

cvp = c(0, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 1.0)
cvp_values = quantile(var_country$cv, na.rm = TRUE, probs = cvp)
cvp_names = names(cvp_values)
cvp_names[1] = "Min"
cvp_names[length(cvp_names)] = "Max"

report.table(data.frame(cvp_names, cvp_values), c('Percentile', 'CV'), "CV Percentiles")

```

```{r esg-var-percs2, fig.cap = "Density of ESG indicator CV", fig.margin = FALSE, fig.width = 6}

plot = var_country %>% ggplot(aes(x = cv)) +
  geom_histogram(aes(y = ..density..),    # histogram chart
                 alpha = 0.8,
                 position = 'identity',
                 bins = 50) +
  viridis::scale_fill_viridis(discrete=TRUE) +
  viridis::scale_color_viridis(discrete=TRUE) +
  hrbrthemes::theme_ipsum() +
  theme(
    legend.position = "none",
    panel.spacing = unit(0.1, "lines"),
    strip.text.x = element_text(size = 8),
    panel.grid = element_blank()
  ) +
  xlab("Coefficient of variation") +
  ylab("K-density") +
  geom_density(alpha = .2, fill = "#CCFFFF") +     # add density char
  geom_vline(aes(xintercept = median(cv, na.rm = TRUE)), color = "#FF3333", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cvp_values["5%"]), color = "#3399FF", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cvp_values["95%"]), color = "#3399FF", linetype = "dashed", size = 1)

report.graphic(plot, "esg-indicator-cv")
```

One important feature of Table \@ref(tab:esg-var-percs) and Figure \@ref(fig:esg-var-percs2) is the presence of a substantial number
of zero values, which account for
`r sprintf("%.1f%%", nrow(filter(var_country, cv==0)) *100 / nrow(filter(var_country, !is.na(cv))))`
of the total number of data points. This occurs when all observations for a country/indicator series are exactly the same
value. Examples include highly stable statistics such as land area, renewable freshwater resources, and electricity mix (coal, nuclear, etc). In some countries, health indicators such as AIDS prevalence also exhibit no variation. Other cases
may be greater cause for concern. For instance, the fact that the "net migration" statistics for Madagascar, Comoros, and Burkina Faso have not changed over 15 years seems dubious at first glance.


A similar-looking spike in distribution occurs around the 95th percentile. This spike is dominated by a handful
of indicators that include observations for only 2-4 years, such as "Population living in areas of elevation below
5 meters," such that any variation at all results in a significant standard deviation and thus an outsized CV value.

Figure \@ref(fig:esg-var) shows the distribution of mean CV values (at the indicator level)
for the ESG indicator dataset. Again, the the red
line marks the median, while the blue lines mark the 5th and 95th percentiles.
The mean CV value may help economists assess whether gap filling is practical for a given indicator.
Indicators with low mean CVs may be especially good candidates for gap filling except for the small number of countries with
individually high CV values. Indicators with relatively high mean CVs may not be good candidates.

```{r esg-var, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Density of mean ESG indicator CV", fig.margin = FALSE, fig.width = 6}

cvimean_values = quantile(var_ind$cv, na.rm = TRUE, probs = c(0.05, 0.95))

plot = var_ind %>% ggplot(aes(x = cv)) +
  geom_histogram(aes(y = ..density..),    # histogram chart
                 alpha = 0.8,
                 position = 'identity',
                 bins = 15) +
  viridis::scale_fill_viridis(discrete=TRUE) +
  viridis::scale_color_viridis(discrete=TRUE) +
  hrbrthemes::theme_ipsum() +
  theme(
    legend.position = "none",
    panel.spacing = unit(0.1, "lines"),
    strip.text.x = element_text(size = 8),
    panel.grid = element_blank()
  ) +
  xlab("Mean Coefficient of variation") +
  ylab("K-density") +
  geom_density(alpha = .2, fill = "#CCFFFF") +     # add density char
  geom_vline(aes(xintercept = median(cv, na.rm = TRUE)), color = "#FF3333", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cvimean_values["5%"]), color = "#3399FF", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cvimean_values["95%"]), color = "#3399FF", linetype = "dashed", size = 1)

report.graphic(plot, "mean-esg-indicator-cv")

```

Table \@ref(tab:t-lh-cv) shows the indicators whose mean CVs fall below the 5th percentile or above the 95th percentile:

```{r t-lh-cv, echo = FALSE}

report.table(t_lh2, c(lvarn, hvarn), "ESG indicators with lowest and highest volatility",
             pdf=list(widths=list(12, 12)))
```

## Implications for Gap Filling ##

While gaps in ESG data follow several patterns as described in the [Coverage Analysis](#coverage) chapter, the most
significant issue is likely indicators whose most recent value (MRV) is 2 years out of date or more, as investment
decisions are made in the context of current and near future conditions. The CV analysis in this chapter and summarized
in Figures \@ref(fig:esg-var-percs2) and \@ref(fig:esg-var) suggest that (normalized) indicator values fall within a fairly narrow range with few outliers, and that the majority of values are below 1.0.
In [several probability fields](https://en.wikipedia.org/wiki/Coefficient_of_variation#Applications)
values below 1.0 are considered low-variance. While it's not clear whether this benchmark applies in the context of some
or all ESG indicators, it provides a starting point for thinking--conservatively--about the potential for using
statistical methods for filling gaps in ESG data. Again, it is beyond the scope of this analysis to suggest
specific gap filling techniques, but stipulating that gap filling will be more robust where indicator variance is low, we can start to think about the impacts of doing so.

```{r, echo=FALSE}
# would be great to replace this with live code...
nCountry = select(var_country, iso3c) %>% unique() %>% nrow()
nSeries = select(var_country, indicator) %>% unique() %>% nrow()
nPairs = nCountry * nSeries
n2018Values = 7659
n2018ValuesB = 8559

select_estimates <- data.frame(
  cv_thresh=c(0.551, 0.688, 0.873, 0.873),
  years_to_extrapolate=c(1, 1, 1, 2),
  result=sprintf('%0.f', c(0.35, 0.39, 0.43, 0.50) * 100))

```

For instance, in this analysis there are `r nCountry` countries and `r nSeries` indicators for a total of
`r nPairs` country/indicator pairs. `r n2018Values` country/indicator series have an MRV value of 2018 or later.
So we can say that the year 2018 has coverage of *`r nf(n2018Values)` / `r nf(nPairs)`,* or
`r sprintf('%.0f%%', n2018Values*100/nPairs)`. To
estimate the potential for gap filling techniques, let's say that we can confidently extrapolate
or estimate an additional year of data from observed values beyond the current MRV for any country/indicator series with
a CV value below the 25th percentile of `r cvp_values['25%']`. Again, while we don't know exactly how the estimate
would be obtained,  we can assume the effect would be to increase the MRV year by 1, raising 
2016 to 2017, 2017 to 2018, etc. The resulting new estimates mean that we now have `r nf(n2018ValuesB)` country/indicator series with at least a 2018 MRV value, which raises our coverage to
*`r nf(n2018ValuesB)` / `r nf(nPairs)`* or
`r sprintf('%.0f%%', n2018ValuesB*100/nPairs)`. One could vary the study year, CV threshold and number of years
to extrapolate in various combinations, as shown in Table \@ref(tab:cv-estimates).

```{r cv-estimates}

report.table(select_estimates,
             c('CV Threshold', 'Years To Extrapolate', 'Resulting Coverage (%)'),
             'Effects of selected CV thresholds and lengths of imputation on coverage in 2018',
             align='r')
#knitr::kable(select_estimates, "html",
#  col.names=,
#  row.names=FALSE, table.attr='class="striped"',
#  align=rep('r', 3),
#  caption='Effects of selected CV thresholds and lengths of imputation on coverage in 2018')
```

```{r tool_paragraph, results='asis'}

if( knitr::is_html_output() ) {
  insert = "below"
} else {
  insert = sprintf("[available here](%s)", metadata$imputation_tool_url)
}

cat("Alternatively, the tool", insert, " provides an interactive way to test
various combinations of assumptions and see both the aggregate impacts and the effects on
individual indicators. Area under the curve (AUC) is equivalent to the percent coverage for any
set of assumptions (100% being \"perfect\" coverage of a value for every indicator and country).")

```

```{r cv-interactive, fig.width = 14, eval=knitr::is_html_output()}
knitr::include_app(metadata$imputation_tool_url, height = "800px")
```



