# Household surveys availability

```{r setuphh, echo=FALSE, message=FALSE, warning=FALSE}
#--------- execute script of analysis
source("R/load_data.R")
source("R/esg_hhAvailability.R")

```

```{r hh-indi, echo = FALSE}
a <- mtd %>% 
  filter(str_detect(source_instrument, "HHS")) %>% 
  select(sector, input_name) 
colnames(a) <- c("Sector", "Indicator")
knitr::kable(a, format = 'html', table.attr = 'class=\"margintab marginnote\"',
        caption = 'Indicators from household surveys')
```

Among the `r mtd %>% summarise(n_distinct(cetsid))` ESG indicators, `r mtd %>% filter(str_detect(source_instrument, "HHS")) %>% summarise(n_distinct(input_name))` of them come from surveys. Some of them come from labor force surveys, others come from living standard measurement surveys, and others from independent surveys or enterprise surveys. Each type of survey has characteristics that the others do not have, but, in general, surveys are characterized by being an expensive instrument, whose feasibility depends on budget. It requires planning, extensive execution, training of personal, data cleaning, and many other factors that makes their frequency unstable. 

In this section, we explore the availability of household surveys---which are the underlying data of global poverty estimates---as a case study to understand the general complexity of the indicators that are calculated from microdata. Though not all the conclusions and insights from this case study apply to all surveys, a proper analysis of household surveys analysis can be adequate to inductively infer the context of other ESG indicators that use other types of surveys as their underlying data. 


## Household surveys for Global Poverty monitoring

The biggest repository of household surveys for Global poverty monitoring is known as [PovcalNet](http://iresearch.worldbank.org/PovcalNet/home.aspx) It is an online toohosted by the World Bank that was developed with the purpose of public replication of the World Bank’s poverty measures at the international poverty line (IPL). PovcalNet contains poverty estimates from more than 1,600 household surveys spanning 164 economies and over 40 years, from 1977 to 2017. 

The household survey data are typically collected by national statistical offices in each country, and then compiled, processed, and vetted for inclusion in the Global Monitoring Database (GMD), the World Bank’s repository of multitopic income and expenditure household surveys used to monitor global poverty and shared prosperity. Selected variables have been harmonized to the extent possible such that levels and trends in poverty and other key sociodemographic attributes can be reasonably compared across and within countries over time. The GMD’s harmonized microdata are used in PovcalNet and the GDSP (World Bank 2018).

## Gaps between surveys

```{r hh-p-time, echo=FALSE, fig.cap = "Production of household surveys over time", fig.margin = TRUE}
g_hhp
```
Poverty rates are based on one of two types of welfare aggregate measure, income or consumption. Countries choose their measure of preference, but, it general, poor countries tent to measure poverty with consumption surveys, whereas rich countries tent to do it with income surveys.  Consumption surveys have a wide range of question and are thus more time-consuming to answer. Also, they are more expensive than income surveys and thus they are carried over about every five years. In contrast, income surveys are less expensive and mostly available every year. So, if a country measures poverty using a consumption-based survey, poverty and inequality measures will be available every five years in average. In the early 70s and 80s, the gap between surveys was so wide that poverty measures were available about every ten years. Today. The number of surveys available is larger and the year gaps between them are shorter. Figure \@ref(fig:hh-p-time) shows the production of surveys over time by region. As expected, the most recent years show a decay in the production, not because no surveys are beeing conducted, but because they have not been released yet. Overall it seems that all regions are doing very well on improving the number of household surveys available to measure poverty.

```{r hh-nc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Number of countries per survey gap", fig.margin = TRUE, fig.width = 5, fig.asp = .75}
g_hh
```

At the country level, however, the frequency of household surveys varies widely.  Figure \@ref(fig:hh-nc) shows the number of countries in each combination of lustrum and number of years gap between surveys. 
