# Coverage Analysis {#coverage}

```{r setupcoverage, echo=FALSE, message=FALSE, warning=FALSE}
#--------- execute script of analysis
source("R/load_data.R")
source("R/esg_availability.R")
#source("R/esg_variability.R")
```


The availability of ESG indicators by country is dynamic over time. As shown in Figure \@ref(fig:ciavailability), not all indicators are available in all countries neither they are available in all years. The bottom left of the chart shows the less available indicators, whereas the top right of the chart shows the most available indicators. From a different perspective, the availability of indicators goes from red to blue. Several insights are drawn from this chart. 

```{r ciavailability, fig.width = 10, fig.height = 10, fig.cap = "Number of countries per indicator over time"}
pg1
```

## Low coverage

```{r echo = FALSE}
tr_low <- 100
d3 <- d2  %>%
  mutate(max = max(mean),
         less = (mean < tr_low)) %>%
  summarise(sless = sum(less),
            max = mean(max),
            s   = sless/max)
```

Some indicators have very low availability over time. In general, `r d3[["sless"]]` (`r round(100*d3[["s"]],2)`%) indicators are available for less than `r tr_low` countries in average over time. Yet, not all indicators with low coverage have the same characteristics. For instance, 

* __Indicator is available since the beginning of the 80's__. Indicators like _Gross National Savings_ or _Adult literacy rate_ are available since 1980. However, they are available YYY countries in average each year. 
* __Indicator is available every few years__. Indicators like _Internal renewable freshwater resources_ or _Net migration_ are available for more than 160 countries but only every five years. Something similar happens to  _PM 2.5 air pollution_, but after 2010 this indicator becomes available every year rather than every five years. 
* __High availability only in recent years.__ Some indicators like _Terrestrial and marine protected areas_ or _Strength of legal rights index_ have a high country coverage but low time span. The former started on 2016 with a coverage of 195 countries, whereas the latter started on 2013 with a coverage of 173 countries. Another indicator that started relatively late is _Scientific and technical journal articles_. It started on 2003 for about 180 countries but has a three-year gap as it is only available until 2016. 

## Remarkable improvement over time

There are a few indicators that show a remarkable improvement in country coverage over time. Despite their early appearance in the beginning of the 1980's, indicators that come from household surveys such as the _Poverty headcounts_ or _Gini index_ have shown a dramatic improvement in country coverage in the beginning of the 2000's.  The coverage improved from about five countries in the 80's to more than 70 countries per year in the decade of the 2000's.  However, these indicators suffer from a compilation lag that requires more than two years from time or recollection to the time of publication. 

Other indicators, especially the ones related to the governance component.  have improved the frequency of publication. For instance, _Governance effectiveness_, _Political stability and absence of violence_, and _Voice and accountability_ were collected every three years at the decade of 1990, but increased their frequency to every year since the 2000. 

## Remain stable over time but with middle coverage

All the indicators with a stable yellow  vertical line are characterized by being stable over time but with a country coverage of about 120 countries per year. Among those, we could find _Prevalence of HIV in adult population_, _non-residents patents application_, or _Physicians per 1,000 people_. 

## High coverage and sudden decline

With the exception of a few indicators like _Ease of doing business index_ or _Total tax and contribution rate_, most recent value (MRV) of most indicators is two or more years old. However, some indicators with remarkable good coverage in the previous decade suddenly declined in the last five years. For instance, _CO2 emissions_ or _Renewable electricity output_ had a yearly coverage of more than 180 countries until 2014/15 and thereafter it suddenly declined to one country, Sudan, for _CO2 emissions_ or zero countries for _Renewable electricity output_. A similar story happened to _Net energy imports_, _Electricity production from nuclear sources_, or _Hospital beds_. However, this set of indicators did not have such a good coverage as _CO2 emissions_ before the decline. 

## high coverage 

Finally, the indicators at the right side of the chart present the best country coverage over time. They are characterized by their intense blue color. For instance, _Population density_, _Adjusted savings: mineral depletion_, or _Land area_ have a remarkable and stable high country coverage. 


