# Explanation Framework Analysis {#explanations}

```{r setupexplanation, echo=FALSE, message=FALSE, warning=FALSE}
source("R/esg_availability.R")
source("R/esg_explanations_summary.R")

# to correctly described the metadata database, we need the original copy before any transforms
original_metadata = read.csv('./data/esg_metadata.csv')
```

The previous chapter sought to provide insights into the patterns of coverage gaps currently found
in ESG data. This section seeks to provide further insight into why these gaps occur.

To better understand the nature of data gaps we consulted data management experts at the World Bank to
develop a list of explanatory factors which are described below. These "explanations" are defined such that they can be used
to consistently "tag" indicators to which they apply. They are particularly relevant to time series statistics.
Using metadata in the World Bank API and interviews with members
of the data management team, we examined each indicator through the lens of these explanations and recorded the results
as dummy (i.e., yes/no) values in a metadata database. It is possible that multiple explanations apply to any given indicator.

The results of this tagging exercise were used to build a database along with other relevant metadata.
The final database assembled for analysis has a total of `r ncol(original_metadata)` variables for 
`r nrow(original_metadata)` indicators.

The discusion below describes each of the explanations in detail. Each description includes a coverage heat
map from the previous section which identifies the indicators "tagged" underneath it, showing how the explanation
may impact gaps in data coverage.


## Description of Explanations

### Ideal Case {-}

The World Bank API defines 217 countries and economies. Indicators that had observations for 2018 or later
for at least 196 countries (90%) were considered "non gap" indicators and were not tagged with
any further explanations.

```{r hm-expl, fig.width = 8, fig.cap = "Ideal Case"}
hm_expl(d1a, ideal, "Ideal case")
```

### Dropped or Deprecated (A) {-}

While most major World Bank time series databases are updated on regular schedules, it is common for individual indicators
and occasionally entire databases to be discontinued. This can occur if the original provider or program discontinues
support for the data, or if policies, programs, or methodologies evolve such that the data are no longer relevant.
In these cases, the time series is no longer updated with recent observations, and the series becomes increasingly obsolete.
Deprecated indicators in the WDI are typically removed and are available only in the [WDI Archives][wdi-archives].
Accordingly, it is not possible to provide a heatmap of these indicators.

 ```{r hm-expla}
 ex <- c('SM.POP.NETM', 'SH.DTH.COMM.ZS', 'ER.H2O.INTR.K3', 'ER.H2O.INTR.PC',
  'NW.NCA.SACO.TO', 'NW.NCA.SAOI.TO', 'NW.NCA.SAGA.TO',
  'SH.STA.DIAB.ZS', 'ER.H2O.FWTL.ZS', 'EN.POP.EL5M.ZS', 'EN.MAM.THRD.NO', 'EN.CLC.MDAT.ZS', 'WBL')
 
gsub('\\$', '\\\\$', knitr::kable(mtd %>% filter(expl_a==TRUE) %>% select(input_name,cetsid),
  "html", escape=FALSE,
  col.names=c('Indicator', 'Code'), table.attr='class="striped"',
  caption="Dropped or Deprecated") )
  
 ```

**Definition:** the indicator only exists in a discontinued or "archived" database, including:

* [WDI Archives][wdi-archives] (API source=2)
* [Africa Development Indicators][adi] (API source=11)

**Notes:** indicators tagged with this explanation are considered inactive and remaining explanations are not considered.


### No Longer Updated (B) {-}

In other cases, an indicator may remain in an active database but fall out of active maintenance. This may be the case as a precursor
to Explanation A above, or it may be attributable to lapsed oversight, i.e., there is no longer a person or team responsible
for the indicator (e.g., program termination, department reorganization, personnel changes). In these cases as well, the time series
becomes increasingly obsolete as it is no longer updated. Some indicators that fit this definition however,
may simply be subject to very long update cycles of 4-5 years or more, meaning that they may still be actively
maintained.

```{r hm-explb, fig.width = 8, fig.cap = "No Longer Updated"}
hm_expl(d1a, expl_b, "No Longer Updated")
```

**Definition:** the indicator has no observations for any country for the last four years (2015-2018)

**Notes:** indicators tagged with this explanation are considered inactive and remaining explanations are not considered.

### Structural Lags (C) {-}

Some indicators by their nature may be especially time consuming to produce. For instance, if an indicator relies on
administrative data or other underlying raw information, it cannot be calculated until that source is published. These
source dependencies may be subject to their own production schedules and timeliness issues. The effect may be that an indicator
may only be available significantly later than the time period of its observations.
These delays may be compounded if multiple dependencies are involved. In Figure \@ref(fig:hm-explc)
indicators that embody this characteristic exhibit low or non-existent coverage on the right side of the heat map.

A different kind of structural lag manifests itself as consistent but relatively low country coverage
compared to the "ideal case" scenario. For instance, an indicator may have consistent temporal coverage, but for
only 85% or less of possible countries, even though the indicator may be relevant to nearly all countries.
Low country coverage may be explained by resource constraints, in that data are more costly to collect in countries
where capacity is limited. It may also be that certain economies (for instance, low population or island states) were
considered out of scope by data producers. In Figure \@ref(fig:hm-explc) indicators that embody this characteristic
exhibit consistent coverage (i.e., uniformly light orange or yellow) along the horizontal axis.


```{r hm-explc, fig.width = 8, fig.cap = "Structural Lags"}
hm_expl(d1a, expl_c, "Structural Lags")
```

**Definition:** classification under this explanation was made through consultations with subject matter experts.

### Curation Lags (D) {-}

In many cases the World Bank is a distributor of indicators it obtains from dozens of external parties.
There is some level of "curation" overhead costs to obtain, collate, validate and finalize data from these independent
sources, and this overhead may itself affect data availability compared to that of the original sources.
This may be especially true where the production cycle of the original indicator varies from that of the Bank's curation team;
for instance, an indicator published monthly by the original source may only be updated quarterly or annually by the World Bank.

```{r hm-expld, fig.width = 8, fig.cap = "Curation Lags"}
hm_expl(d1a, expl_d, "Curation Lags")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.

### Licensing Constraints (E) {-}

In some cases, the World Bank distributes indicators under legal agreements that affect its availability. The most common
cases are licenses that stipulate that the Bank cannot distribute the latest version of a dataset to protect that
dataset's value in cases where providers sell the data commercially. Thus, while more recent data may exist, the Bank is
legally prevented from distributing it for a period of time, which can result in an availability gap.

```{r hm-exple, fig.width = 8, fig.cap = "Licensing Constraints"}
hm_expl(d1a, expl_e, "Licensing Constraints")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.

### Survey Dependencies (F) {-}

Some indicators are calculated from underlying [microdata](https://stats.oecd.org/glossary/detail.asp?ID=1656) obtained from
household-level or firm-level surveys conducted in country. The survey process constitutes a unique kind of structural lag.
Surveys themselves are costly, and are frequently implemented irregularly or according to schedules that vary significantly by country.
As a result, data availability may vary not only by time but by country.

```{r hm-explf, fig.width = 8, fig.cap = "Survey Dependencies"}
hm_expl(d1a, expl_f, "Survey Dependencies")
```

**Definition:** classification under this explanation was made through consultations with the data curation team.


### Limited Relevance (G) {-}

Some indicators may not be relevant to or may not be interesting in the context of certain groups of economies. For example,
very small economies may not have significant natural resource endowments or particular kinds of economic activity for particular
indicators to be meaningful. Indicators designed to measure use of, say, forests, mineral deposits, or levels of trade may
be assumed to essentially be zero or "too small to measure."
In a similar vein, some indicators may not be relevant to industrialized
economies, such as prevalence of certain diseases (thought to be eradicated), literacy rates, or foreign aid.

```{r hm-explg, fig.width = 8, fig.cap = "Limited Relevance"}
hm_expl(d1a, expl_g, "Limited Relevance")
```

**Definition:** this explanation was considered for high-income "rich" countries (using the WBG income classification) and small economies,
defined as those with populations under 120,000. This population threshold represents a "natural" breakpoint at which population jumps by nearly
40 percent, clearly differentiating economies above and below it. For both "rich" and "small" economies we identified a set of indicators
for which at least 80% of economies in the group had no observations at all, as this might suggest a deliberate
decision on the part of data producers to exclude the group.
From these results we removed indicators that we still considered potentially relevant to rich/small
economies despite the lack of available data.


## Findings 
```{r echo=FALSE}
n_total_all = nrow(mtd)
n_total_wbg = nrow(filter(mtd, wbgv1==TRUE))
n_struc_all = nrow(filter(mtd, expl_c==TRUE))
n_struc_wbg = nrow(filter(mtd, expl_c==TRUE, wbgv1==TRUE))
n_surv_all = nrow(filter(mtd, expl_f==TRUE))
n_surv_wbg = nrow(filter(mtd, expl_f==TRUE, wbgv1==TRUE))
n_stale_all = nrow(filter(mtd, expl_b==TRUE))
n_stale_wbg = nrow(filter(mtd, expl_b==TRUE, wbgv1==TRUE))
n_retired_all = nrow(filter(mtd, expl_a==TRUE))
n_retired_wbg = nrow(filter(mtd, expl_a==TRUE, wbgv1==TRUE))
n_curation_all = nrow(filter(mtd, expl_d==TRUE))
```


Figure \@ref(fig:smry-exlp) shows the results of the explanation framework 
disaggregated primarily by explanation, and secondarily by primary source 
(WBG and all external sources).

```{r smry-exlp, fig.width = 12, fig.height = 5.5, fig.cap = "Explanations for data gaps, by source"}
smry_expl <- smry_expl +
  labs(
    # title = "Explanations for gaps: All indicators vs WBG dataset",
    subtitle = "Most data gaps are due to structural lag and lack of survey data. 
    This is true both when considering all indicators, or the subset of indicators
    contained in the WBG dataset.",
    y = "Frequency",
    x = "",
    fill = "Data producer"
  )

plotly::ggplotly(smry_expl, tooltip = "text") %>% 
  plotly::style(hoveron = "fill") %>%
  plotly::layout(legend = list(orientation = 'h', y = -0.1))

```



Overall, the "Structural Lags" explanation is by far the most dominant cause 
of gaps in ESG data, where it is a contributing factor for `r n_struc_all` of `r n_total_all` indicators 
overall, and `r n_struc_wbg` of `r n_total_wbg` indicators in the WBG's ESG dataset. This frequency likely
reflects the prevalence of traditional statistical indicators currently used in ESG, many of 
which happen to rely on inter-agency cooperation and other intrinsically time consuming 
modalities to produce comparable data for all countries.

"Survey Dependencies" is the second most significant explanation, contributing 
to gaps for `r n_surv_all` of `r n_total_all` indicators overall, and
`r n_surv_wbg` of `r n_total_wbg` indicators in the WBG's dataset.
"No Longer Updated" is the third most significant explanation, accounting for `r n_stale_all` of `r n_total_all`
indicators overall, and `r n_stale_wbg` of `r n_total_wbg` indicators in the WBG's dataset. "Deprecated" is close
behind, accounting for `r n_retired_all` of `r n_total_all` indicators overall, but is not a factor in the WBG's dataset.

Of the remaining factors, "Licensing" and "Limited Relevance" are the most significant, and in the WBG dataset these
are equal to "No Longer Updated" in terms of prevalence. The least significant explanation in both datasets was
"Curation Lags," which accounted for just `r n_curation_all` indicators in both datasets.

Figure \@ref(fig:smry-exlp-sctr) is similar to Figure \@ref(fig:smry-exlp) but disaggregates by explanation and sector.


```{r smry-exlp-sctr, fig.width = 12, fig.height = 5.5, fig.cap = "Explanations for data gaps, by sector"}
smry_expl_sctr <- smry_expl_sctr +
  labs(
    # title = "Explanations for gaps by sector: All indicators vs WBG dataset",
    subtitle = "Not only most data gaps are due to structural lags, it is one of
    the most frequent explanation for all three ESG sectors.",
    y = "Frequency",
    x = "",
    fill = "Sector"
  )

plotly::ggplotly(smry_expl_sctr, tooltip = "text") %>% 
  plotly::style(hoveron = "fill") %>%
  plotly::layout(legend = list(orientation = 'h', y = -0.1))
```

This provides a slightly different perspective. Looking at the top 2 explanatory factors, indicators in the
social and governance pillars are overly represented within "Structural Lags" and "Survey Dependencies," suggesting
that efforts in these areas would disproportionaately improve these pillars. By comparison, nearly all indicators
in the "No Longer Updated" group belong to the environment pillar. The "Licensing" explanation also consists entirely
of indicators in the environment pillar.



[wdi-archives]: https://datacatalog.worldbank.org/dataset/wdi-database-archives
[adi]: https://datacatalog.worldbank.org/dataset/africa-development-indicators
